---
title: "World_insects_estimate"
format: pdf
editor: visual
---

Load libraries and ../Data

```{r, message=FALSE}
library(SpadeR)
library(dplyr)
library(stringr)
library(pairwiseCI)
#library(remotes)
#install_version("CARE1", "1.1.0")
library(CARE1)

## Load Data 

#Microgastrinae 

reared <- read.csv("../Data/Microgastrinae ACG Reared.csv") %>% 
  select(BOLD.code, Frequency)

core <- read.csv("../Data/Microgastrinae ACG Core Malaise.csv") %>% 
  select(BOLD.code, Frequency)

peri <- read.csv("../Data/Microgastrinae ACG Peripheral Malaise.csv") %>% 
  select(BOLD.code, Frequency)

# All insecta core

insc <- read.csv("../Data/Insecta ACG Core Malaise.csv") %>%
  select(BOLD.code, Frequency)

## highlight estimator

estimator <- "    iChao1 (Chiu et al. 2014)"

## adding in the trees

#TA = number of tree species in ACG
TA <- 1500
TA_alt <- 1200

#Tc = number of tree species in CR

TC <- 2500
#Tw = number of tree species on Earth

## estimated tree species

TW <- 73274
TW_CI <- c(73271, 73276)

TW_2 <- 89147

## described number of tree species 

TW_d <- 60065

### Number of Saturniid moths

SA <- 106
SW <- 4535

```

# Approach 1: Calculate insect richness worldwide using the CARE estimator for Microgastrinae

```{r}
### Total number of BINS

reared %>% nrow()

core %>% nrow()

peri %>% nrow()

###
```

## Step 1: Combining Malaise and Reared samples of Microgastrinae and estimating the true richness of ACG Microgastrinae

```{r}

full_micro_table <- reared %>% 
  rename(reared = Frequency) %>% 
  full_join(rename(core, core = Frequency)) %>% 
  full_join(rename(peri, peri = Frequency)) %>% 
  mutate(reared = ifelse(is.na(reared), 0, 1)) %>% 
  mutate(core = ifelse(is.na(core), 0, 1)) %>% 
  mutate(peri = ifelse(is.na(peri), 0, 1))

head(full_micro_table)

nrow(full_micro_table)

# Transform observed ../Data to CARE1 input format
micro_care <- as.record(full_micro_table[,-1])

micro_care

# Main step to obtain population size estimates
CARE1.print(micro_care)

# Print the CARE1 input format

## n1 = reared = 886
## n2 = core = 388
## n3 = peripheral = 576

## Using Nhat-1

Nhat1 <- round(estN(micro_care, method = "LSC", se = TRUE, nboot = 200))

micro_all_estimate <- Nhat1[1]
micro_all_CI <- c(Nhat1[3], Nhat1[4])
```

## Step 2. Estimating the undersampling ratio for Core Microgastrinae.

```{r}
## proportion of core malaise of the ichao total 

proportion_core <- nrow(core)/micro_all_estimate

proportion_core

CI_proportion_core <- nrow(core)/micro_all_CI

CI_proportion_core
```

## Step 3 Calculate total number of insects in ACG

```{r}
## Total number of insecta bins in ACG core malaise
n_insects_core <- nrow(insc)

n_insects_core
```

```{r}
total_insec_acg <- n_insects_core*(1/proportion_core)

total_insec_acg

CI_total_insect_acg <- n_insects_core*(1/round(CI_proportion_core,3))

n_insects_core*(1/round(CI_proportion_core,3))

n_insects_core*(1/CI_proportion_core)

CI_total_insect_acg
```

## Step 4. Estimating global insect richness – based on trees

```{r}

##### Total insects 

print("Total world insects based on estimated tree richness")

total_insect_world <- (TW*total_insec_acg)/TA 
total_insect_world

CI_total_insect_world <- (TW*CI_total_insect_acg)/TA 

CI_total_insect_world

print("Total world insects based on described trees")

total_insect_world <- (TW_d*total_insec_acg)/TA 
total_insect_world

CI_total_insect_world <- (TW_d*CI_total_insect_acg)/TA 

CI_total_insect_world

##### Total insects 

print("Total world insects based on estimated tree richness")

total_insect_world <- (TW*total_insec_acg)/TA_alt 
total_insect_world

CI_total_insect_world <- (TW*CI_total_insect_acg)/TA_alt 

CI_total_insect_world

print("Total world insects based on described trees")

total_insect_world <- (TW_d*total_insec_acg)/TA_alt 
total_insect_world

CI_total_insect_world <- (TW_d*CI_total_insect_acg)/TA_alt

CI_total_insect_world
```

## Step 4. Estimating global insect richness – based on saturniid moths

```{r}

##### Total insects 

print("Total world insects based on Saturniid moths")

total_insect_world <- (SW*total_insec_acg)/SA 
total_insect_world

CI_total_insect_world <- (SW*CI_total_insect_acg)/SA 

CI_total_insect_world

```


## An alternative estimate

```{r}

##### Total insects 

print("Total world insects based on estimated tree richness")

total_insect_world <- (TW_2*total_insec_acg)/TA 
total_insect_world

CI_total_insect_world <- (TW_2*CI_total_insect_acg)/TA 

CI_total_insect_world


##### Total insects 

print("Total world insects based on estimated tree richness")

total_insect_world <- (TW_2*total_insec_acg)/TA_alt 
total_insect_world

CI_total_insect_world <- (TW_2*CI_total_insect_acg)/TA_alt 

CI_total_insect_world

```

