---
title: "World_insects_estimate"
format: pdf
editor: visual
---

Load libraries and ../Data

```{r, message=FALSE}
library(SpadeR)
library(dplyr)
library(stringr)
library(pairwiseCI)
#library(remotes)
#install_version("CARE1", "1.1.0")
library(CARE1)

## Load Data 

#Microgastrinae 

reared <- read.csv("../Data/Microgastrinae ACG Reared.csv") %>% 
  select(BOLD.code, Frequency)

core <- read.csv("../Data/Microgastrinae ACG Core Malaise.csv") %>% 
  select(BOLD.code, Frequency)

peri <- read.csv("../Data/Microgastrinae ACG Peripheral Malaise.csv") %>% 
  select(BOLD.code, Frequency)

# All insecta core

insc <- read.csv("../Data/Insecta ACG Core Malaise.csv") %>%
  select(BOLD.code, Frequency)

## highlight estimator

estimator <- "    iChao1 (Chiu et al. 2014)"

## adding in the trees

#TA = number of tree species in ACG
TA <- 1500
TA_alt <- 1200

#Tc = number of tree species in CR

TC <- 2500
#Tw = number of tree species on Earth

## estimated tree species

TW <- 73274
TW_CI <- c(73271, 73276)

## described number of tree species 

TW_d <- 60065

### Number of Saturniid moths

SA <- 106
SW <- 4535

```

# Approach 1: Calculate insect richness worldwide using the CARE estimator for Microgastrinae

```{r}
### Total number of BINS

reared %>% nrow()

core %>% nrow()

peri %>% nrow()

###
```

## Step 1: Combining Malaise and Reared samples of Microgastrinae and estimating the true richness of ACG Microgastrinae

```{r}

full_micro_table <- reared %>% 
  rename(reared = Frequency) %>% 
  full_join(rename(core, core = Frequency)) %>% 
  full_join(rename(peri, peri = Frequency)) %>% 
  mutate(reared = ifelse(is.na(reared), 0, 1)) %>% 
  mutate(core = ifelse(is.na(core), 0, 1)) %>% 
  mutate(peri = ifelse(is.na(peri), 0, 1))

head(full_micro_table)

nrow(full_micro_table)

# Transform observed ../Data to CARE1 input format
micro_care <- as.record(full_micro_table[,-1])

micro_care

# Main step to obtain population size estimates
CARE1.print(micro_care)

# Print the CARE1 input format

## n1 = reared = 886
## n2 = core = 388
## n3 = peripheral = 576

## Using Nhat-1

Nhat1 <- round(estN(micro_care, method = "LSC", se = TRUE, nboot = 200))

micro_all_estimate <- Nhat1[1]
micro_all_CI <- c(Nhat1[3], Nhat1[4])
```

## Step 2. Estimating the undersampling ratio for Core Microgastrinae.

```{r}
## proportion of core malaise of the ichao total 

proportion_core <- nrow(core)/micro_all_estimate

proportion_core

CI_proportion_core <- nrow(core)/micro_all_CI

CI_proportion_core
```

## Step 3 Calculate total number of insects in ACG

```{r}
## Total number of insecta bins in ACG core malaise
n_insects_core <- nrow(insc)

n_insects_core
```

```{r}
total_insec_acg <- n_insects_core*(1/proportion_core)

total_insec_acg

CI_total_insect_acg <- n_insects_core*(1/round(CI_proportion_core,3))

n_insects_core*(1/round(CI_proportion_core,3))

n_insects_core*(1/CI_proportion_core)

CI_total_insect_acg
```

## Step 4. Estimating global insect richness – based on trees

```{r}

##### Total insects 

print("Total world insects based on estimated tree richness")

total_insect_world <- (TW*total_insec_acg)/TA 
total_insect_world

CI_total_insect_world <- (TW*CI_total_insect_acg)/TA 

CI_total_insect_world

print("Total world insects based on described trees")

total_insect_world <- (TW_d*total_insec_acg)/TA 
total_insect_world

CI_total_insect_world <- (TW_d*CI_total_insect_acg)/TA 

CI_total_insect_world

##### Total insects 

print("Total world insects based on estimated tree richness")

total_insect_world <- (TW*total_insec_acg)/TA_alt 
total_insect_world

CI_total_insect_world <- (TW*CI_total_insect_acg)/TA_alt 

CI_total_insect_world

print("Total world insects based on described trees")

total_insect_world <- (TW_d*total_insec_acg)/TA_alt 
total_insect_world

CI_total_insect_world <- (TW_d*CI_total_insect_acg)/TA_alt

CI_total_insect_world
```

## Step 4. Estimating global insect richness – based on saturniid moths

```{r}

##### Total insects 

print("Total world insects based on Saturniid moths")

total_insect_world <- (SW*total_insec_acg)/SA 
total_insect_world

CI_total_insect_world <- (SW*CI_total_insect_acg)/SA 

CI_total_insect_world

```

# Additional Information

For Pooled Microgastrinae CV and Sample Coverage: 1-\[number singletons\]/\[number specimens\].

```{r}
### Combine all microgastrinae

micro_all <- rbind(reared, core, peri) %>% 
  group_by(BOLD.code) %>% 
  summarise(new_freq = sum(Frequency)) %>% 
  select(BOLD.code, new_freq)

## number of pooled micros
nrow(micro_all)

## use chao on microgastrinae

micro_all_estimate <- ChaoSpecies(micro_all$new_freq,"abundance",k=10,conf=0.95)

micro_all_estimate$Species_table[estimator,]
## from spade R
micro_all_estimate$Basic_data_information[c("    CV for entire dataset",
                                            "    Coverage estimate for entire dataset"),]

## based on coverage spreadsheet 

total_bins <- nrow(micro_all)
total_bins

total_singletons_micro_all <- sum(micro_all$new_freq ==1)
total_singletons_micro_all

total_doubletons_micro_all <- sum(micro_all$new_freq ==2)
total_doubletons_micro_all

# total samples
sum(micro_all$new_freq)

coverage_micro_all <- 1-(total_singletons_micro_all/sum(micro_all$new_freq))
coverage_micro_all

CV_micro_all <- sd(micro_all$new_freq)/mean(micro_all$new_freq)
CV_micro_all
```

Sample Coverage For Core Malaise Microgastrinae.

```{r}

## from Spade R
micro_core_estimate <- ChaoSpecies(core$Frequency,"abundance",k=10,conf=0.95)

micro_core_estimate$Basic_data_information[c("    CV for entire dataset",
                                            "    Coverage estimate for entire dataset"),]

## calculating from the spreadsheet

total_bins <- nrow(core)
total_bins

total_singletons_core <- sum(core$Frequency ==1)
total_singletons_core

total_doubletons_core <- sum(core$Frequency ==2)
total_doubletons_core

# total samples
sum(core$Frequency)

coverage_core <- 1-(total_singletons_core/sum(core$Frequency))
coverage_core

CV_core <- sd(core$Frequency)/mean(core$Frequency)
CV_core


```

Sample Coverage For Core Insecta

```{r}
## from spade R

insc_acg_estimate <- readRDS("../Data/insecta_acg_chao.rds")

insc_acg_estimate$Basic_data_information[c("    CV for entire dataset",
                                            "    Coverage estimate for entire dataset"),]

insecta <- read.csv("../Data/Insecta ACG Core Malaise.csv")

## from the spreadsheet

total_bins <- nrow(insecta)
total_bins

total_singletons_insecta <- sum(insecta$Frequency ==1)
total_singletons_insecta

total_doubletons_insecta <- sum(insecta$Frequency ==2)
total_doubletons_insecta

coverage_insecta <- 1-(total_singletons_insecta/sum(insecta$Frequency))
coverage_insecta

CV_insecta <- sd(insecta$Frequency)/mean(insecta$Frequency)
CV_insecta


```

## Number of singletons and sample coverage for Core Malaise for each taxonomic level (Braconidae, Ichneumonoidea, and Hymenoptera).

### Braconidae

```{r}

braconidae <- read.csv("../Data/Braconidae ACG Core Malaise.csv")

## from the spreadsheet

total_bins <- nrow(braconidae)
total_bins

total_singletons_braconidae <- sum(braconidae$Frequency ==1)
total_singletons_braconidae

total_doubletons_braconidae <- sum(braconidae$Frequency ==2)
total_doubletons_braconidae

## total number of samples

sum(braconidae$Frequency)

coverage_braconidae <- 1-(total_singletons_braconidae/sum(braconidae$Frequency))
coverage_braconidae

CV_braconidae <- sd(braconidae$Frequency)/mean(braconidae$Frequency)
CV_braconidae


```

### Ichneumonoidea

```{r}

ichneumonoidea <- read.csv("../Data/Ichneumonoidea ACG Core Malaise.csv")

## from the spreadsheet

total_bins <- nrow(ichneumonoidea)
total_bins

total_singletons_ichneumonoidea <- sum(ichneumonoidea$Frequency ==1)
total_singletons_ichneumonoidea

total_doubletons_ichneumonoidea <- sum(ichneumonoidea$Frequency ==2)
total_doubletons_ichneumonoidea

## total number of samples

sum(ichneumonoidea$Frequency)

coverage_ichneumonoidea <- 1-(total_singletons_ichneumonoidea/sum(ichneumonoidea$Frequency))
coverage_ichneumonoidea

CV_ichneumonoidea <- sd(ichneumonoidea$Frequency)/mean(ichneumonoidea$Frequency)
CV_ichneumonoidea


```

### Hymenoptera

```{r}

hymenoptera <- read.csv("../Data/Hymenoptera ACG Core Malaise.csv")

## from the spreadsheet

total_bins <- nrow(hymenoptera)
total_bins

total_singletons_hymenoptera <- sum(hymenoptera$Frequency ==1)
total_singletons_hymenoptera

total_doubletons_hymenoptera <- sum(hymenoptera$Frequency ==2)
total_doubletons_hymenoptera

## total number of samples

sum(hymenoptera$Frequency)

coverage_hymenoptera <- 1-(total_singletons_hymenoptera/sum(hymenoptera$Frequency))
coverage_hymenoptera

CV_hymenoptera <- sd(hymenoptera$Frequency)/mean(hymenoptera$Frequency)
CV_hymenoptera


```
