---
title: "Supplementary analyses"
format: pdf
editor: visual
date: "`r format(Sys.Date(), '%B %d, %Y')`"
---

## Table S1

Using the Macrogastrinae benchmark to estimate ACG and global Braconidae, Ichneumonoidea, and Hymenoptera (following the same approach we have used for Insecta). We have added Table S1 to present these estimates, which we discuss briefly in the Main Text.

```{r, message=FALSE}
library(SpadeR)
library(dplyr)
library(stringr)
library(pairwiseCI)
#library(remotes)
#install_version("CARE1", "1.1.0")
library(CARE1)

## Load Data 

#Microgastrinae 

reared <- read.csv("../Data/Microgastrinae ACG Reared.csv") %>% 
  select(BOLD.code, Frequency)

core <- read.csv("../Data/Microgastrinae ACG Core Malaise.csv") %>% 
  select(BOLD.code, Frequency)

peri <- read.csv("../Data/Microgastrinae ACG Peripheral Malaise.csv") %>% 
  select(BOLD.code, Frequency)

# All insecta core

insc <- read.csv("../Data/Insecta ACG Core Malaise.csv") %>%
  select(BOLD.code, Frequency)

#Braconidae

Brac <- read.csv("../Data/Braconidae ACG Core Malaise.csv") %>%
  select(BOLD.code= BOLD.Code, Frequency)

#Ichneumonoidea

Ichn <- read.csv("../Data/Ichneumonoidea ACG Core Malaise.csv") %>%
  select(BOLD.code, Frequency)

#Hymenoptera

hym <- read.csv("../Data/Hymenoptera ACG Core Malaise.csv") %>% 
  select(BOLD.code= BOLD.Code, Frequency)

## highlight estimator

estimator <- "    iChao1 (Chiu et al. 2014)"

## adding in the trees

#TA = number of tree species in ACG
TA <- 1500
TA_alt <- 1200

#Tc = number of tree species in CR

TC <- 2500
#Tw = number of tree species on Earth

## estimated tree species

TW <- 73274
TW_CI <- c(73271, 73276)

TW_2 <- 89147

## described number of tree species 

TW_d <- 60065

### Number of Saturniid moths

SA <- 106
SW <- 4535

```

# Approach 1: Calculate insect richness worldwide using the CARE estimator for Microgastrinae

```{r}
### Total number of BINS

reared %>% nrow()

core %>% nrow()

peri %>% nrow()

###
```

## Step 1: Combining Malaise and Reared samples of Microgastrinae and estimating the true richness of ACG Microgastrinae

```{r}

full_micro_table <- reared %>% 
  rename(reared = Frequency) %>% 
  full_join(rename(core, core = Frequency)) %>% 
  full_join(rename(peri, peri = Frequency)) %>% 
  mutate(reared = ifelse(is.na(reared), 0, 1)) %>% 
  mutate(core = ifelse(is.na(core), 0, 1)) %>% 
  mutate(peri = ifelse(is.na(peri), 0, 1))

head(full_micro_table)

nrow(full_micro_table)

# Transform observed ../Data to CARE1 input format
micro_care <- as.record(full_micro_table[,-1])

micro_care

# Main step to obtain population size estimates
CARE1.print(micro_care)

# Print the CARE1 input format

## n1 = reared = 886
## n2 = core = 388
## n3 = peripheral = 576

## Using Nhat-1

Nhat1 <- round(estN(micro_care, method = "LSC", se = TRUE, nboot = 200))

micro_all_estimate <- Nhat1[1]
micro_all_CI <- c(Nhat1[3], Nhat1[4])
```

## Step 2. Estimating the undersampling ratio for Core Microgastrinae.

```{r}
## proportion of core malaise of the ichao total 

proportion_core <- nrow(core)/micro_all_estimate

proportion_core

CI_proportion_core <- nrow(core)/micro_all_CI

CI_proportion_core
```

## Step 3 Calculate total number of insects in ACG

```{r}
## Total number of insecta bins in ACG core malaise
n_insects_core <- nrow(insc)

n_insects_core

## Total number of Braconidae bins in ACG core malaise
n_brac_core <- nrow(Brac)

n_brac_core

## Total number of Ichneumonoidea bins in ACG core malaise
n_ichn_core <- nrow(Ichn)

n_ichn_core

## Total number of hymenoptera bins in ACG core malaise
n_hym_core <- nrow(hym)

n_hym_core

```

```{r}

## insecta 
total_insec_acg <- n_insects_core*(1/proportion_core)

total_insec_acg

CI_total_insect_acg <- n_insects_core*(1/round(CI_proportion_core,3))

n_insects_core*(1/round(CI_proportion_core,3))

n_insects_core*(1/CI_proportion_core)

CI_total_insect_acg


## Braconidae 
total_brac_acg <- n_brac_core*(1/proportion_core)

total_brac_acg

CI_total_brac_acg <- n_brac_core*(1/round(CI_proportion_core,3))

n_brac_core*(1/round(CI_proportion_core,3))

n_brac_core*(1/CI_proportion_core)

CI_total_brac_acg

## Ichneumonoidea  
total_ichn_acg <- n_ichn_core*(1/proportion_core)

total_ichn_acg

CI_total_ichn_acg <- n_ichn_core*(1/round(CI_proportion_core,3))

n_ichn_core*(1/round(CI_proportion_core,3))

n_ichn_core*(1/CI_proportion_core)

CI_total_ichn_acg

## Hymenoptera 
total_hym_acg <- n_hym_core*(1/proportion_core)

total_hym_acg

CI_total_hym_acg <- n_hym_core*(1/round(CI_proportion_core,3))

n_hym_core*(1/round(CI_proportion_core,3))

n_hym_core*(1/CI_proportion_core)

CI_total_hym_acg

## compile results

compiled_results <- data.frame(Taxonomic_group = c('Insecta', "Hymenoptera", 
                                                   "Ichneumonoidea", "Braconidae"),
                               total_core = c(n_insects_core, n_hym_core, 
                                              n_ichn_core, n_brac_core),
                               total_acg = c(total_insec_acg, total_hym_acg,
                                            total_ichn_acg, total_brac_acg),
                               CI_low = c(CI_total_insect_acg[1],
                                          CI_total_hym_acg[1],
                                          CI_total_ichn_acg[1],
                                          CI_total_brac_acg[1]),
                               CI_high = c(CI_total_insect_acg[2],
                                          CI_total_hym_acg[2],
                                          CI_total_ichn_acg[2],
                                          CI_total_brac_acg[2]))

compiled_results
```

## Step 4. Estimating global insect richness – based on trees

```{r}

##### Total insects 

print("Total world insects based on estimated tree richness")

total_insect_world <- (TW*total_insec_acg)/TA 
total_insect_world

CI_total_insect_world <- (TW*CI_total_insect_acg)/TA 

CI_total_insect_world


##### Total Hymenoptera 

print("Total world insects based on estimated tree richness")

total_hym_world <- (TW*total_hym_acg)/TA 
total_hym_world

CI_total_hym_world <- (TW*CI_total_hym_acg)/TA 

CI_total_hym_world


##### Total Ichneumonoidea

print("Total world insects based on estimated tree richness")

total_ichn_world <- (TW*total_ichn_acg)/TA 
total_ichn_world

CI_total_ichn_world <- (TW*CI_total_ichn_acg)/TA 

CI_total_ichn_world


##### Total Braconidae 

print("Total world insects based on estimated tree richness")

total_brac_world <- (TW*total_brac_acg)/TA 
total_brac_world

CI_total_brac_world <- (TW*CI_total_brac_acg)/TA 

CI_total_brac_world

## compile results

compiled_results_all <- bind_cols(compiled_results, 
                                  data.frame(total_word = 
                                               c(total_insect_world,
                                                 total_hym_world,
                                                 total_ichn_world,
                                                 total_brac_world),
                                             CI_w_low = c(CI_total_insect_world[1],
                                                          CI_total_hym_world[1],
                                                          CI_total_ichn_world[1],
                                                          CI_total_brac_world[1]),
                                             CI_w_high = c(CI_total_insect_world[2],
                                                          CI_total_hym_world[2],
                                                          CI_total_ichn_world[2],
                                                          CI_total_brac_world[2])))

compiled_results_all

```
